---
title: "Assignment 3"
format: html
editor: visual
---

***How to do it?***:

- Open the Quarto file of this assignment ([link](assignment3.qmd)) in Rstudio.

- Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

- Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

## Problems

\

1.  Create the following data frame

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
df <- data.frame(Rank = c(0, 1, 2, 3), 
                Age = c(28, 34, 29, 42), 
                Name = c("Tom", "Jack", "Steve", "Ricky"))

df
```

2.  Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://bryantstats.github.io/math421/data/us-states.csv).

```{r}
df <- read.csv("https://bryantstats.github.io/math421/data/us-states.csv")
head(df)
```

3.  Show the names of the variables in the data

```{r}
names(df)
```

4.  How many columns and rows the data have?

```{r}
str(df)
```

There are 5 columns and 51,078 rows in the data.



5.  How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?

```{r}
sum(is.na(df))
colSums(is.na(df))
```
There are no missing values in any of the columns in this dataset.

6.  What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.

```{r}
class(df$date)
```
```{r}
df$date <- as.Date(df$date)
```

```{r}
class(df$date)
```


7.  Capitalize the names of all the variables

```{r}
names(df) <- toupper(names(df))
names(df)
```


8.  Find the average number of cases per day. Find the maximum cases a day.

```{r}
# Aggregate cases so one row is one date
daily_cases <- aggregate(CASES ~ DATE, df, sum)
head(daily_cases)
```


```{r}
# Average
mean(daily_cases$CASES)
```

```{r}
# Maximum
max(daily_cases$CASES)
```

9.  How many states are there in the data?

```{r}
unique(df$STATE)
```
```{r}
length(unique(df$STATE))
```
There are 56 states in the data.


10. Create a new variable `weekdays` to store the weekday for each rows.

```{r}
df$WEEKDAYS <- weekdays(df$DATE)
head(df)
```


11. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
- `no_death` if there is no death that day

```{r}
df$DEATH2 <- ifelse(df$DEATHS > 0, 'has_death', 'no_death')
head(df)
```


Find the frequency and relative frequency of `no_death` and `has_death`.

```{r}
# Frequency
table(df$DEATH2)
```
```{r}
# Relative Frequency
prop.table(table(df$DEATH2))
```

12. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)

```{r}
summary(df$DEATHS)
```

13. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
library(dplyr)

Q1 <- 838
Q3 <- 12674

df$DEATH3 <- case_when(df$DEATHS < Q1 ~ 'low_death',
                       df$DEATHS <= Q3 ~ 'mid_death',
                       TRUE ~ 'high_death')

head(df)
```


14. Find the average cases in Rhode Island in 2021

```{r}
df_RI  <- df[df$STATE == 'Rhode Island',]
df_RI_2021  <- df_RI[df_RI$DATE >= '2021-01-01' & df_RI$DATE <= '2021-12-31',]
head(df_RI_2021)
```

```{r}
mean(df_RI_2021$CASES)
```

15. Find the median cases by weekdays in Rhode Island in 2021

```{r}
by(df_RI_2021$CASES, df_RI_2021$WEEKDAYS, median)
```

16. Compare the median cases in Rhode Island in June, July, August and September in 2021.

```{r}
df_RI_June_2021 <- df_RI[df_RI$DATE >= '2021-06-01' & df_RI$DATE <= '2021-06-30',]
df_RI_July_2021 <- df_RI[df_RI$DATE >= '2021-07-01' & df_RI$DATE <= '2021-07-31',]
df_RI_August_2021 <- df_RI[df_RI$DATE >= '2021-08-01' & df_RI$DATE <= '2021-08-31',]
df_RI_September_2021 <- df_RI[df_RI$DATE >= '2021-09-01' & df_RI$DATE <= '2021-09-30',]
```

```{r}
median(df_RI_June_2021$CASES)
median(df_RI_July_2021$CASES)
median(df_RI_August_2021$CASES)
median(df_RI_September_2021$CASES)

```
In this comparison, June has the lowest median cases and September has the highest median cases in 2021.


17. Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when

```{r}
df1 = read.csv("https://bryantstats.github.io/math421/data/titanic.csv")
```

```{r}
# head
head(df1)
```

```{r}
#str
str(df1)
```
```{r}
# names
names(df1)
```
```{r}
# mean, min, max
mean(df1$Age, na.rm = TRUE)
min(df1$Age, na.rm = TRUE)
max(df1$Age, na.rm = TRUE)
```
```{r}
# table
table(df1$Sex)
```
```{r}
# is.na, colSums
sum(is.na(df1))
colSums(is.na(df1))
```
```{r}
# class
class(df1$Fare)
```

```{r}
# cor
cor(df1$Fare, df1$Survived)
```


```{r}
# by

# % of people who survived by class
by(df1$Survived, df1$Pclass, mean)
```


```{r}
# ifelse
df1$traveled_alone <- ifelse(df1$SibSp == 0, 'Yes', 'No')
head(df1)

```


```{r}
# case_when

library(dplyr)


df1$age_bins <- case_when(df1$Age < 18 ~ 'Child',
                       df1$Age <= 65 ~ 'Adult',
                       is.na(df1$Age) ~ 'Unknown',
                       TRUE ~ 'Elderly')

head(df1)
```


18. In the dataset in #17, practice the follows. You can reuse the code of 17.

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories

```{r}
summary(df1$Fare)
```


```{r}
# Create categorical variable from a continuous variable
library(dplyr)

Q1 <- 7.91
Q3 <- 31

df1$fare_bins <- case_when(df1$Fare < Q1 ~ 'Low',
                       df1$Fare <= Q3 ~ 'Medium',
                       TRUE ~ 'High')

head(df1)
```


```{r}
# From one categorical variable, create a new categorical variable with fewer categories

df1$FirstClass <- ifelse(df1$Pclass == 1, 'Yes', 'No') 

head(df1)

```


If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](https://bryantstats.github.io/math421/data/titanic.csv)
