
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 8: Visualization - Making a bar race"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment8.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install two packages `gganimate` and `gifski` then restart Rstudio. Using the `Adult Census Income` data, make an animation using `geom_point` and `transition_states`.

```{r, eval=FALSE}
# Install Packages
install.packages("gganimate")
install.packages("gifski")
```

```{r}
library(gganimate)
library(gifski)
library(tidyverse)
df <- read_csv("https://bryantstats.github.io/math421/data/adult_census.csv")
str(df)
```

```{r}
df |> ggplot(aes(x = education.num, y = hours.per.week, color = race)) + 
  geom_point() +
  transition_states(sex) +
  labs(title = 'sex: {closest_state}')
```


2. Using the `Adult Census Income` data, make an animation using `geom_bar` and `transition_states`.

```{r}
df |> ggplot(aes(x = income, fill = sex)) + 
  geom_bar(position = 'fill') +
  transition_states(race) +
  labs(title = 'race: {closest_state}')
```


3. Use the WHO's dataset at [this link](https://covid19.who.int/WHO-COVID-19-global-data.csv). Make a top-10 bar race by months between countries on the number of deaths by Covid 19 in 2021.

```{r}
df1 <- read_csv("https://bryantstats.github.io/math421/data/WHO-COVID-19-global-data.csv")
str(df1)
```

```{r}
# Data
library(lubridate)
df1_data <- df1 |> 
  filter(year(Date_reported) == 2021) |> 
  group_by(month = month(Date_reported, label = TRUE, abbr = FALSE), Country) |> 
  summarise(total_new_deaths = sum(New_deaths))

df1_data_2 <- df1_data |> 
  group_by(month) |> 
  mutate(rank=rank(-total_new_deaths))

df1_data_3 <- df1_data_2 |>  filter(rank <= 10)
```

```{r}
# Bar race
p1 <- df1_data_3 |> 
  ggplot(aes(x = rank, y = total_new_deaths, group = Country, fill = Country, label = Country)) + 
  geom_col() + 
  geom_text(aes(y = total_new_deaths, 
                  label = Country), 
              hjust = 1.4) + 
    coord_flip(clip = "off", 
               expand = FALSE)+
    scale_x_reverse()+
    labs(title='{closest_state}', 
         y='Total New Deaths', x='', fill='')+
    theme(plot.title = 
          element_text(hjust=1,size=22),
          axis.ticks.y=element_blank(),
          axis.text.y=element_blank(),
          legend.position = "bottom") + 
    transition_states(month) +
    ease_aes("cubic-in-out")
animate(p1, nframes=400)
```


4. Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (https://covidtracking.com/data/download/all-states-history.csv) but you should make a different bar racev from ones in the slides. 

```{r}
df2 <- read_csv("https://bryantstats.github.io/math421/data/all-states-history.csv")
str(df2)
```


```{r}
# Data
df2_1 <- df2 |> filter(deathIncrease > 0, hospitalizedIncrease > 0, positiveIncrease > 0)

df2_2 <- df2_1 |> group_by(week = week(date), state) |> summarise(mean = mean((hospitalized/positive) * 100))
df2_3 <- df2_2|> group_by(week) |> mutate(rank = rank(-mean))
df2_4 <- df2_3 |> filter(rank <= 10)

# Bar race
plot <- df2_4 |>  ggplot(aes(x = rank, y = mean, group = state, fill = state, label = state)) + geom_col()+
    geom_text(aes(y = mean, label = state), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Week {closest_state}', x = '', y = 'Hosiptalization Rate (%)', fill = 'state') +
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(week)+
    ease_aes("cubic-in-out")
animate(plot, nframes = 400)
```

