
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.

```{r}
library(tidyverse)

df <- read_csv("https://bryantstats.github.io/math421/data/adult_census_missing.csv")
df
```

```{r}
colSums(is.na(df))
```



2. Using function `aggr` the VIM package to plot the number of NAs for each column. 

```{r}
library(VIM)

aggr(df)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data.

```{r}
table(df$age)
```
```{r}
table(df$workclass)
```
```{r}
table(df$education)
```
```{r}
table(df$education.num)
```

```{r}
table(df$marital.status)
```

```{r}
table(df$occupation)
```
```{r}
table(df$relationship)
```
```{r}
table(df$race)
```

```{r}
table(df$sex)
```
```{r}
table(df$capital.gain)
```
```{r}
table(df$capital.loss)
```

```{r}
table(df$hours.per.week)
```

```{r}
table(df$native.country)
```

```{r}
table(df$income)
```

**The  other forms of NAs are ?, Unknown, and Not Known.**


4. Replace all the forms of missing values found to NA

```{r}
df <- replace(df, df == 'Unknown' | 
                  df == '?' | 
                  df == 'Not Known', NA)
```


5. Replot the number of NAs for each column. 

```{r}
colSums(is.na(df))
```


```{r}
aggr(df)
```


6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?

```{r}
df1 <- drop_na(df)
df1
```

```{r}
colSums(is.na(df1))
```
**29,865 rows are left after removing the ones with NAs.**

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 

```{r}
df2 <- df |> fill(age, workclass, education, marital.status, occupation, relationship, sex, capital.gain, native.country, .direction = 'updown')
df2
```

```{r}
colSums(is.na(df2))
```

8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
# Get classes of each column
sapply(df, class)
```


```{r}
# Numeric
median_age <- median(df$age, na.rm = TRUE)
median_capital_gain <- median(df$capital.gain, na.rm = TRUE)


# Categorical
majority_workclass <- names(which.max(table(df$workclass)))
majority_education <- names(which.max(table(df$education)))
majority_marital_status <- names(which.max(table(df$marital.status)))
majority_occupation <- names(which.max(table(df$occupation)))
majority_relationship <- names(which.max(table(df$relationship)))
majority_sex <- names(which.max(table(df$sex)))
majority_native_country <- names(which.max(table(df$native.country)))

```


```{r}
df3 <- df
```

```{r}
df3$age <- replace_na(df3$age, median_age)
df3$capital.gain <- replace_na(df3$age, median_capital_gain)
df3$workclass <- replace_na(df3$age, majority_workclass)
df3$education <- replace_na(df3$age, majority_education)
df3$marital.status <- replace_na(df3$age, majority_marital_status)
df3$occupation <- replace_na(df3$age, majority_occupation)
df3$relationship <- replace_na(df3$age, majority_relationship)
df3$sex <- replace_na(df3$age, majority_sex)
df3$native.country <- replace_na(df3$age, majority_native_country)

df3
```

```{r}
colSums(is.na(df3))
```

