
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

This assignment works with the [IMDB Top 1000 data](../data/imdb_top_1000.csv).  Find out more information about this data at [this link](https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows).  Import the data and answer the following questions. 

1. List all the names of the columns of the data

```{r}
library(tidyverse)
df <- read_csv("C:\\Users\\student\\Downloads\\imdb_top_1000.csv")
df
```

```{r}
names(df)
```

2. Which movies have the highest money earned (Gross)?

```{r}
df |> arrange(-Gross)
```

**The top 5 movies with the highest money earned are Star Wars: Episode VII - The Force Awakens, Avengers: Endgame, Avatar, Avengers: Infinity War, and Titanic.**

3. What is the lowest rating (IMDB_Rating)?  List five movies have this lowest rating.

```{r}
df |> arrange(IMDB_Rating)
```

**The lowest rating is 7.6. Five movies with this rating are Dark Waters, Searching, Once Upon a Time... in Hollywood, Nelyubov, and The Florida Project.**

4. Which year have the most number of movies released in the list?  What is the total of money earned on that year?

```{r}
df |> 
  group_by(Released_Year) |> 
  summarise(count = n(), sum(Gross, na.rm = TRUE)) |> 
  arrange(-count)
```

**2014 had the most movies released and grossed $2,755,629,221 in total.**

5. What is the total money earned per movies on average?

```{r}
df |> summarise(avg_gross = mean(Gross, na.rm = TRUE))
```

**$68,034,751**


6. Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9.

```{r}
# Average number of votes by year
df |> 
  group_by(Released_Year) |> 
  summarise(mean(No_of_Votes))
```

```{r}
# Average number of votes of movies that have IMDB rating greater than 9
df |> 
  filter(IMDB_Rating > 9) |> 
  summarise(mean(No_of_Votes))
```


7. Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile. 

```{r}
summary(df$No_of_Votes)
```


```{r}
Q2 <- 138549
Q3 <- 374161

df |> 
  filter(Released_Year == 2020 & No_of_Votes >= Q2 & No_of_Votes <= Q3) |> 
  summarise(avg_meta_score = mean(Meta_score, na.rm = TRUE))
```


8. (Optional - Challenging). The current `Runtime` variable is not a numeric. Use the `str_remove` function to remove ` min` from the variables then use `as.numeric` to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding `use="complete.obs"` in the `cor` function to ignore the missing values).  

```{r}
table(df$Runtime)
```

```{r}
# Remove `min`
df <- df |> mutate(Runtime = str_remove_all(Runtime, ' min'))
table(df$Runtime)
```

```{r}
# Convert to numeric
df$Runtime <- as.numeric(df$Runtime)
class(df$Runtime)
```

```{r}
# Average running time in the 2010s
df |> 
  filter(Released_Year >= 2010 & Released_Year < 2020) |> 
  summarise(mean(Runtime))
```

```{r}
# Correlation between running time and IMDB rating
cor(df$Runtime, df$IMDB_Rating, use="complete.obs")
```

9. We can use `select_if` to select columns satisfying a condition and use `summarise_if` to do calculation on columns satisfying a condition.  Try the follows to understand these functions. 

```{r,eval=FALSE}
# Select only character columns
df %>% select_if(is.character)

# Calculate the median of all numeric columns
df %>% summarise_if(is.numeric, mean, na.rm=TRUE)
```


10. Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.

- select
- filter
- mutate
- summarise
- arrange
- count
- count + arrange
- filter + count + arrange
- group_by + summarise
- filter + group_by + summarise
- filter + group_by + summarise + arrange

```{r}
# select: deselect the `Poster_Link` column
df |> select(-Poster_Link)
```

**Deselecting this column makes the dataframe look less cluttered.**

```{r}
# filter: filter the dataframe to only rated R movies
df |> filter(Certificate == 'R')
```

**There are only 146 movies that are rated R.**

```{r}
# mutate: make an indicator column for whether the movie was released in the 2000s or not
df <- df |> mutate(Movie_2000s = ifelse(Released_Year >= 2000, 'Y', 'N'))
```


```{r}
table(df$Movie_2000s)
```

**Most movies were released before the 2000's, but only by 28 movies.**

```{r}
# summarise: find the max runtime of a movie.
df |> 
  summarise(max(Runtime))
```

**The maximum runtime of a movie is 321 minutes.**

```{r}
# arrange: what is the highest metascore a movie can have? What are 5 movies with this score?
df |> arrange(-Meta_score)
```

**The highest Meta score is 100. Five movies with this score are The Godfather, Casablanca, Rear Window, Lawrence of Arabia, and Vertigo.**

```{r}
# count: who are the directors of the movies? Who are the first 3 in alphabetical order?
df |> count(Director)

```

**Aamir Khan, Aaron Sorkin, and Abdellatif Kechiche**

```{r}
# count + arrange: which director directed the most movies?
df |> 
  count(Director) |> 
  arrange(-n)

```

**Alfred Hitchcock directed the most movies.**

```{r}
# filter + count + arrange: in what year did Leonardo DiCaprio star in the most movies?
df |> 
  filter(Star1 == "Leonardo DiCaprio" | Star2 == "Leonardo DiCaprio" | Star3 == "Leonardo DiCaprio" | Star4 == "Leonardo DiCaprio") |>
  count(Released_Year) |> 
  arrange(-n)
```

**Leonardo DiCaprio starred in the most movies in 2006 in 2010 releases.**

```{r}
# group_by + summarise: What is the average IMDB rating of movies per year? List the first 3 release years and the average rating.
df |> 
  group_by(Released_Year) |> 
  summarise(avg_IMDB_rating = mean(IMDB_Rating, na.rm = TRUE))
```

**1920: 8.1, 1921: 8.3, 1922: 7.9**

```{r}
# filter + group_by + summarise: of the movies Tom Hanks stars in, what is the average IMDB rating by director? Name the last 3 alphabetically.
df |> 
  filter(Star1 == "Tom Hanks" | Star2 == "Tom Hanks" | Star3 == "Tom Hanks" | Star4 == "Tom Hanks") |> 
  group_by(Director) |> 
  summarise(mean(IMDB_Rating, na.rm = TRUE))

```

**Ron Howard: 7.6, Sam Mendes: 7.7, Steven Spielberg: 8.1**


```{r}
# filter + group_by + summarise + arrange: of the movies that are only in the comedy genre, what certificate has the highest average Meta score? 
df |> 
  filter(Genre == 'Comedy') |> 
  group_by(Certificate) |> 
  summarise(avg_meta_score = mean(Meta_score, na.rm = TRUE)) |> 
  arrange(-avg_meta_score)
```

**PG has the highest average meta score at 93.**